A Temporal Logic Based Framework for Intrusion Detection

### 2023-09-07 Paper/main.tex
update

\textbf{Calculus.} According to~\cite{barringer2004rule}, the $evaluate$, $update$ and $value$ functions are defined a priori for all operators except for the
rule application, for which get generated based on the definition of rules in the specification. The definitions of these functions on the different primitive operators are the following ($op\text{ can be }\lor , \land, \rightarrow $):
%evaluate
{\allowdisplaybreaks
\begin{align*}
    evaluate(\underline{\text{true}},\alpha_i) & =\ \underline{\text{true}} \\
    evaluate(\underline{\text{false}},\alpha_i) & =\ \underline{\text{false}} \\ 
    evaluate(exp,\alpha_i) & =\ \text{value of } exp\text{ in } \alpha_i \\
    evaluate(F_1\ op\ F_2,\alpha_i) & =\  evaluate(F_1,\alpha_i)\ op\ evaluate(F_2,\alpha_i)\\
    evaluate(\neg{F},\alpha_i) & =\ \neg{evaluate(F,\alpha_i)}  \\
    evaluate(\bigcirc F,\alpha_i) & =\ update(F,\alpha_i)  \\
    evaluate(F_1 \cdot F_2,\alpha_i) & =
    \begin{cases}
        evaluate(F_1,\alpha_i)\cdot F_2 \quad \text{if } value(F_1)=\ \underline{\text{false}} \\
        evaluate(F_1,\alpha_i)\cdot F_2 \lor evaluate(F_2,\alpha_i)\quad \text{else} 
    \end{cases} \\ \\
%\end{align*}
%\noindent\begin{minipage}[l]{\textwidth}
   % \raggedright
    %\begin{align*}
        value(\underline{\text{true}}) & = \underline{\text{true}} \\
        value(\underline{\text{false}}) & = \underline{\text{false}} \\ 
        value(\text{exp}) & = \underline{\text{false}} \\
        value(F_1\ \text{op}\ F_2) & = value(F_1)\ \text{op}\ value(F_2) \\
        value(\neg{F}) & = \neg{value(F)}  \\
        value(\bigcirc F) & =
        \begin{cases}
            F & \text{if at the beginning of a trace} \\
            \underline{\text{false}} & \text{if at the end of a trace} \\
        \end{cases}\\
        value(F_1 \cdot F_2,\alpha_i) & = value(F_1)\land value(F_2) \\ \\
   % \end{align*}
%\end{minipage}
%\hfill
%\begin{minipage}[r]{\textwidth}
   % \raggedleft
    %\begin{align*}
        update(\underline{\text{true}},\alpha_i) & = \underline{\text{true}} \\
        update(\underline{\text{false}},\alpha_i) & = \underline{\text{false}} \\ 
        update(\text{exp},\alpha_i) & = \text{exp}\\
        update(F_1\ \text{op}\ F_2,\alpha_i) & = update(F_1,\alpha_i)\ \text{op}\ update(F_2,\alpha_i) \\
        update(\neg{F},\alpha_i) & = \neg{update(F,\alpha_i)}  \\
        update(\bigcirc F,\alpha_i) & = update(F,\alpha_i)  \\
        update(F_1 \cdot F_2,\alpha_i) & = update(F_1,\alpha_i)\cdot F_2
    \end{align*}
    }
%\end{minipage}
Let us introduce an example to better understand the algorithm. Given the following rules 
\begin{align*}
    &\underline{\text{max}}\ \mathtt{Always}(\underline{\text{Form}}\ F)  = F \land \bigcirc \mathtt{Always}(F) \\
    &\underline{\text{min}}\ \mathtt{EvTimedLogout}(\underline{\text{string }} k, \underline{\text{double }} t, \underline{\text{double }} \delta) = (\text{time} - t \leq \delta) \\
    &\quad\quad\quad\quad \land ((action = \mathtt{logout} \land userId = k) \lor \bigcirc \mathtt{EvTimedLogout}(k, t, \delta)) \\
    &\underline{\text{mon}}\ M_3  = \mathtt{Always}((action = \mathtt{login}) \rightarrow \mathtt{EvTimedLogout}(userId,time, 100))
\end{align*}
and a trace $\sigma=\alpha_1,\alpha_2$, where $\alpha_1=\{userId:\mathtt{"Lori"},\ action:\mathtt{login},\ time:\mathtt{17.0}\}$ and $\alpha_1=\{userId:\mathtt{"Lori"},\ action:\mathtt{logout},\ time:\mathtt{150.0}\}$, the initial formula $M_3$ gets modified. At time $1$, when $\alpha_1$ is analyzed, $M_3$ changes as follows:
\begin{align*}
    &evaluate(M_3,\alpha_1)  = \mathtt{EvTimedLogout("Lori",17.0, 100)} \\ 
    &\quad\quad\quad\quad \land\mathtt{Always}((action = \mathtt{login}) \rightarrow \mathtt{EvTimedLogout}(userId,time, 100))
\end{align*}
At the second event the predicate $(time-t \leq \delta)$ gets instantiated to $(150.0-17.0 \leq 100)$ which is false; thus the whole formula becomes false, indicating that the two-event trace violates the property $M_3$.

>>>>

\textbf{Calculus.} According to~\cite{barringer2004rule}, the $evaluate$, $update$ and $value$ functions are defined a priori for all operators except for the
rule application, for which get generated based on the definition of rules in the specification. The definitions of these functions on the different primitive operators (except the previous operator $\odot$) are the following ($op\text{ can be }\lor , \land, \rightarrow $):
%evaluate
{\allowdisplaybreaks
\begin{align*}
    evaluate(\underline{\text{true}},\alpha_i) & =\ \underline{\text{true}} \\
    evaluate(\underline{\text{false}},\alpha_i) & =\ \underline{\text{false}} \\ 
    evaluate(exp,\alpha_i) & =\ \text{value of } exp\text{ in } \alpha_i \\
    evaluate(F_1\ op\ F_2,\alpha_i) & =\  evaluate(F_1,\alpha_i)\ op\ evaluate(F_2,\alpha_i)\\
    evaluate(\neg{F},\alpha_i) & =\ \neg{evaluate(F,\alpha_i)}  \\
    evaluate(\bigcirc F,\alpha_i) & =\ update(F,\alpha_i)  \\
    evaluate(F_1 \cdot F_2,\alpha_i) & =
    \begin{cases}
        evaluate(F_1,\alpha_i)\cdot F_2 \quad \text{if } value(F_1)=\ \underline{\text{false}} \\
        evaluate(F_1,\alpha_i)\cdot F_2 \lor evaluate(F_2,\alpha_i)\quad \text{else} 
    \end{cases} \\ \\
%\end{align*}
%\noindent\begin{minipage}[l]{\textwidth}
   % \raggedright
    %\begin{align*}
        value(\underline{\text{true}}) & = \underline{\text{true}} \\
        value(\underline{\text{false}}) & = \underline{\text{false}} \\ 
        value(\text{exp}) & = \underline{\text{false}} \\
        value(F_1\ \text{op}\ F_2) & = value(F_1)\ \text{op}\ value(F_2) \\
        value(\neg{F}) & = \neg{value(F)}  \\
        value(\bigcirc F) & =
        \begin{cases}
            F & \text{if at the beginning of a trace} \\
            \underline{\text{false}} & \text{if at the end of a trace} \\
        \end{cases}\\
        value(F_1 \cdot F_2,\alpha_i) & = value(F_1)\land value(F_2) \\ 
        value(\mathtt{R}(F_1,\ldots,F_n)) & =
        \begin{cases}
            \underline{\text{true}} & \text{if }\mathtt{R}\text{ is } \underline{\text{max}}\\
            \underline{\text{false}} & \text{if }\mathtt{R}\text{ is } \underline{\text{min}}\\
        \end{cases}\\
        value(F_1 \cdot F_2,\alpha_i) & = value(F_1)\land value(F_2) \\ \\
   % \end{align*}
%\end{minipage}
%\hfill
%\begin{minipage}[r]{\textwidth}
   % \raggedleft
    %\begin{align*}
        update(\underline{\text{true}},\alpha_i) & = \underline{\text{true}} \\
        update(\underline{\text{false}},\alpha_i) & = \underline{\text{false}} \\ 
        update(\text{exp},\alpha_i) & = \text{exp}\\
        update(F_1\ \text{op}\ F_2,\alpha_i) & = update(F_1,\alpha_i)\ \text{op}\ update(F_2,\alpha_i) \\
        update(\neg{F},\alpha_i) & = \neg{update(F,\alpha_i)}  \\
        update(\bigcirc F,\alpha_i) & = update(F,\alpha_i)  \\
        update(F_1 \cdot F_2,\alpha_i) & = update(F_1,\alpha_i)\cdot F_2
    \end{align*}
    }
Focusing of \textit{Future Time Operators}, we have:
\begin{align*}
    evaluate(\mathtt{Next}(F),s)&=evaluate(\bigcirc F,s) \\
    evaluate(\mathtt{Always}(F),s)&=evaluate(F\land\bigcirc\mathtt{Always}(F),s) \\
    evaluate(\mathtt{Eventually}(F),s)&=evaluate(F\lor\bigcirc\mathtt{Eventually}(F),s) \\
    evaluate(\mathtt{Until}(F_1,F_2),s)&=evaluate(F_2\lor(F_1\land\bigcirc\mathtt{Until}(F_1,F_2)),s) \\ \\
    update(\mathtt{Next}(F),s)&=\mathtt{Next}(update(F,s)) \\
    update(\mathtt{Always}(F),s)&=\mathtt{Always}(update(F,s)) \\
    update(\mathtt{Eventually}(F),s)&=\mathtt{Eventually}(update(F,s)) \\
    update(\mathtt{Until}(F_1,F_2),s)&=\mathtt{Until}(update(F_1,s),update(F_2,s))
\end{align*}
while, for the \textit{Past Time Operators} the $\odot$ operator is removed and instead a rule $\mathtt{Previous}$ is introduced; more details can be found in~\cite{barringer2004program}.

%\end{minipage}
Let us introduce an example to better understand the algorithm. Given the following rules 
\begin{align*}
    &\underline{\text{max}}\ \mathtt{Always}(\underline{\text{Form}}\ F)  = F \land \bigcirc \mathtt{Always}(F) \\
    &\underline{\text{min}}\ \mathtt{EvTimedLogout}(\underline{\text{string }} k, \underline{\text{double }} t, \underline{\text{double }} \delta) = (\text{time} - t \leq \delta) \\
    &\quad\quad\quad\quad \land ((action = \mathtt{logout} \land userId = k) \lor \bigcirc \mathtt{EvTimedLogout}(k, t, \delta)) \\
    &\underline{\text{mon}}\ M_3  = \mathtt{Always}((action = \mathtt{login}) \rightarrow \mathtt{EvTimedLogout}(userId,time, 100))
\end{align*}
and a trace $\sigma=\alpha_1,\alpha_2$, where $\alpha_1=\{userId:\mathtt{"Lori"},\ action:\mathtt{login},\ time:\mathtt{17.0}\}$ and $\alpha_1=\{userId:\mathtt{"Lori"},\ action:\mathtt{logout},\ time:\mathtt{150.0}\}$, the initial formula $M_3$ gets modified. At time $1$, when $\alpha_1$ is analyzed, $M_3$ changes as follows:
\begin{align*}
    &evaluate(M_3,\alpha_1)  = \mathtt{EvTimedLogout("Lori",17.0, 100)} \\ 
    &\quad\quad\quad\quad \land\mathtt{Always}((action = \mathtt{login}) \rightarrow \mathtt{EvTimedLogout}(userId,time, 100))
\end{align*}
At the second event the predicate $(time-t \leq \delta)$ gets instantiated to $(150.0-17.0 \leq 100)$ which is false; thus the whole formula becomes false, indicating that the two-event trace violates the property $M_3$.
